---
title: "Exploring Great Duck Island Through Data"
author: "Eleanor and Wriley"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(sf)
library(forcats)
```

## 1. Introduction

In this project, we will use existing data about the seabirds breeding on Great Duck Island to explore their ecologies and distributions. Much of this data was collected for long-term monitoring or for student projects that have been completed. We hope that by re-visiting some of this data from a higher-level and data-focused perspective, we will be able to tease out some new information about the island that has yet to be discovered. We are particularly interested in exploring the spatial distribution of the seabird colonies on GDI and how they have changed over time. The first part of our project will focus on visualizing how the gull colony on the south end of the island has changed since 1999, allowing us to develop spatial visualization skills in R. Next, we hope to bring in several years' worth of chick-check data containing chick growth metrics. This will doubtless require a great deal of data tidying, but should yield some interesting visualizations and insights. Doing this will also allow us to consider a best practice for the future collection of this type of data. Later, we may want to tidy and analyze other data including band return data, petrel habitat surveys, gull GPS-tracker data, and more. An ideal outcome of this project would be a tidier and more accessible data storage system for Great Duck Island that would help further researchers collect and organize data in a more productive way. 

The data that will be used in this project have come from two decades of student research on Great Duck Island. The vast majority of it will be taken from the GDI data storage system created by Addison Gruber '22. This system consists of organized files on the island computer and shapefiles in the Great Duck Island group on COA's ArcGIS Online organization. Most of the data that we will use have spatial elements--observations are usually connected to a nest, burrow, or plot on the island. Some examples of quantitative variables that may be present include clutch sizes, bird weights, bird wing chord lengths, numbers of burrows per survey plot. Some examples of categorical variables may include habitat types (forest, meadow, etc), bird band codes, bird species, or flag numbers (in this case categorical because it represents a unique identifier rather than a value).

Great Duck Island is an important seabird colony in the Gulf of Maine--it is the largest Leach's Storm-Petrel colony in the lower 48 United States, and one of the only colonies in the Gulf where gulls are increasing instead of declining. Seabird populations worldwide are of increasing population concern, and large, long-term data-sets on these difficult-to-study birds are rare and hard to obtain. Great Duck Island has a wonderful body of research collected by numerous students across dozens of different projects, some short-term and some long-term. Unfortunately, as students have come and gone and graduated, little follow-up has occurred on data from past years. We hope that considering Great Duck's data holistically, rather than by working on a single specific project, we will be able to discover some interesting trends and patterns that are not yet clear to us, and improve our understanding of this amazing piece of COA's campus. We have a goldmine of data collected on Great Duck. There are aspects of this data that have been well analyzed, and others that are relatively untouched. Our hope for this project is to create successful visualizations and comprehensive analysis from this long term data. We also hope to continue the work of Addison, and continue to organize data from Great Duck. Finally, we plan to create additional metadata where possible, and create a central location from which future students can access this valuable data from the last 25 years.


## 2. Data

A bare-bones example of a shapefile of gull nests mapped on the south end:
```{r gull shape file, echo = FALSE}
gullshape <- read_sf("data/gulls2021")
head(gullshape)
plot(gullshape)
```

An example of raw data from the 2022 petrel survey:

```{r petrel survey data, echo = FALSE}
petrelsurvey <- read_csv("data/islandestimatefinal.csv")
head(petrelsurvey)
```

An example of some petrel band return data from 2022 (which birds were found or banded in which burrow and their morphometrics):
```{r petrel bands, echo = FALSE}
petrelbands <- read_csv("data/longfidelity.csv")
head(petrelbands)
```


## 3. Data analysis plan

Much of our data analysis will involve mapping spatial data, so an early part of our data analysis will include learning to work with this sort of data. We hope to give ourselves a primer in using the r packages "leaflet" and "sf." We have a continuous series of maps from 1999 - 2022 that show the location of each nest. We hope to create a series of exploratory series of data visualizations showing the changing distribution of gull nests that breed on the south end of Great Duck. If we are able to, we would like create an animation demonstrating the changes in distribution. The main variables of interest here in these spatial visualizations are location within the south end colony and year. As an interesting supplement to this, we will also create a bar graph demonstrating island wide shifts in distribution. Gulls have nested at 4 different sub colonies. While the total number of gulls has changed only slightly in the last 25 years, their distribution between these four sub colonies has changhed dramatically.

Another long term data set we have is from tower count observations. Each morning, the Great Duck Team stands in the lighthouse at 0700 and counts all birds in sight. This data set also goes back 20+ years. We would like to create a series of graphs demonstrating the changes in the abundance of birds that we have counted. The variables of interest here are species, count, and year. 

We will create a series of visualizations looking at chick check data. Chick check is where a team of students selects a sample of nests that they visit each day to record weight and survival status. This allows us to look at growth curves and survival rates of chicks. We have a tidy version of this data from 2022 that we will definitely visualize. There are many variables of interest in this 2022 data: habitat of nest, clutch size, hatch sequence, growth rates and survival status are the ones we will explore for this project. We also have data ranging back 20 or so years. These data are in disparate formats. If we have the time, we will begin tidying these data so that we can compare them. There is a lot of potential for comparing data from across the years, but there is also a lot of work that would need to be done to do this. If we succeeded, we could compare over all fledging success and average growth curves across the years.

An interesting visualization of 2022 petrel survey data, where the number of burrows per survey plot is mapped over longitude and latitude. The bimodal distribution of burrows across latitude is very interesting to us. 

Plotting the latitude and longitudes of each subplot (so that each subplot is a point) shows how well we covered the island with our survey--it creates a clear map of Great Duck. 


```{r petrel exploratory visualization, echo = FALSE}
petrelcoords <- read_csv("data/subplotscoords.csv")
ggplot(petrelcoords, aes(x = latitude, y = Total))+
  geom_point()
ggplot(petrelcoords, aes(x = longitude, y = latitude))+
  geom_point()
  
```

Sample distribution of survey plot populations from the 2022 petrel survey:
```{r petrel summary statistics, echo = FALSE}

petrelsurvey %>% 
  filter(!is.na(Subplot)) %>%
  group_by(Plot) %>%
  summarize(burrows_obs = sum(Total),n_subplots=n()) %>% 
  ggplot(aes(x = (burrows_obs*0.46)))+
  geom_histogram(binwidth = 1)+
  xlab("Observed number of active burrows per 50x50 meter plot")

```

An example of a visualization from 2022 gull data, looking at fledging success by habitat:

```{r reading gull data, echo = FALSE}
chickcheck2022 <- read_csv("data/chickchecktidy2022.csv")
fledging2022 <- read_csv("data/fledging_final2022.csv")
```

```{r fledging success barplot in ggplot, echo = FALSE}

#create standard error function
se <- function(x) sd(x)/sqrt(length(x))
#Eleanor check that this is doing what it's supposed to

fledging2022 %>% 
  group_by(habitat) %>% 
  summarise(mean = mean(fledging_success), se = se(fledging_success)) %>%
  ggplot(aes(x = habitat, y= mean, fill = habitat))+
  geom_col()+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2,
                 position=position_dodge(.9))+
  scale_fill_viridis_d()+
  labs(caption = "Error Bars Denote Standard Error", title = " Mean # of Chicks Fledged by Habitat")+
  ylab("Mean number of chicks fledged")

fledging2022 %>% 
  group_by(habitat) %>% 
  summarise(median = median(fledging_success), se = se(fledging_success)) %>%
  ggplot(aes(x = habitat, y= median, fill = habitat))+
  geom_col()+
  geom_errorbar(aes(ymin=median-se, ymax=median+se), width=.2,
                 position=position_dodge(.9))+
  scale_fill_viridis_d()+
  labs(caption = "Error Bars Denote Standard Error", title = "Median # of chicks fledged by habitat")+
  ylab("Median number of chicks fledged")

```

```{r gull exploratory visualization, echo = FALSE}
ggplot(fledging2022, aes(x = fledging_success))+
  geom_bar(fill = "gray" , color = "black")+
  facet_wrap(~habitat) +
  theme_bw()+
  labs(title = "Count of nests that fledged 0 - 3 chicks by habitat")
```






